---
title: "STAT625_Project"
author: "Brian Zhang, Vi Mai, Xinyu Zhou (Anna), Ziyan Zhao"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r}
library(lmerTest)
library(car)
library(ggplot2)
rm(list = ls())
walmart <- read.csv("Walmart.csv")
head(walmart)
```

### Data Preprocessing 

Since dates are strings, they must be converted to parsed and converted to days. Use days since the first day rather than the actual date to make computation easier. 
```{r}
# Convert the dates from character strings into days since the first date 
asDate_result <- as.Date(walmart$Date, "%d-%m-%Y")
first_date <- min(asDate_result)
days_elapsed <- asDate_result-first_date
walmart["Days_since"] <- days_elapsed
head(walmart)


```

### Initial Diagnostic Plots 
```{r}
# Comparing the need for a log transformation on Weekly Sales 
full_mod_no_log <- lm(Weekly_Sales ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart)
plot(full_mod_no_log,
     col = rgb(red = 0, green = 0, blue = 0, alpha = 0.25), 
     pch = 16
)

full_mod_w_log <- lm(log(Weekly_Sales) ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart)
plot(full_mod_w_log,
     col = rgb(red = 0, green = 0, blue = 0, alpha = 0.25), 
     pch = 16
)
```

```{r}
# Fit a mixed model, look at residual plots
mixed_effect_mod <- lmer(Weekly_Sales ~ (1|Store) + Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data=walmart)
plot(mixed_effect_mod)

mixed_effect_mod_log <- lmer(log(Weekly_Sales) ~ (1|Store) + Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data=walmart)
plot(mixed_effect_mod_log)

mixed_effect_mod_sqrt<- lmer(sqrt(Weekly_Sales) ~ (1|Store) + Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data=walmart)
plot(mixed_effect_mod_sqrt)
```

No transformation on Weekly Sales and sqrt(Weekly Sales) show a megaphone effect, nonconstant variance. However, the log transformation on Weekly Sales seems to get rid of the megaphone appearance of the residuals, indicating constant variance. Thus, the log transformation may be the most appropriate. 

```{r}
# QQ plots for different models 
qqPlot(residuals(full_mod_no_log))
qqPlot(residuals(full_mod_w_log))
qqPlot(residuals(mixed_effect_mod))
qqPlot(residuals(mixed_effect_mod_log))

# qq plot for random effects 
qqPlot(ranef(mixed_effect_mod_log)$Store$`(Intercept)`)
```

The full model without the log transform shows residuals that are most close to a normal distribution. However, all show deviations at lower normal quantiles and higher normal quantiles. When comparing the mixed effect on store on log(Weekly Sales) versus a mixed effect on store on Weekly Sales, the one with the log transformation appears to be more close to a normal distribution. 



### Scatterplot with variabels vs log(Weekly_Sales)
```{r}
# Looking at effect of Days and Store on log(Weekly Sales)
plot(log(Weekly_Sales) ~ Days_since + Store + Temperature + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart, 
     col = alpha(walmart$Store, 0.5),  # makes color dependent on store 
     pch = 16,
     cex = 0.6)
```

```{r}
# Scatterplot matrix - uninterprettable
pairs(Weekly_Sales ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart)
```


### Lag Regression 

#### Lag Regression for Holiday
```{r}
# Create lag variable on Holiday
unique_days <- unique(walmart$Days_since)
holiday_flag <- c()
log_average_sales <- c()
for(day in unique_days){
  holiday_flag <- append(holiday_flag, (walmart$Holiday_Flag[walmart$Days_since == day][1]))
  log_average_sales <- append(log_average_sales, mean(log(walmart$Weekly_Sales[walmart$Days_since == day])))
}



ccf_vals <- ccf(holiday_flag, log_average_sales, 4)
ccf_vals
```

```{r}

#143
```


```{r}

x<- data.frame()
for(i in 1:45){
  temp <- subset(walmart, walmart['Store'] == i)
  acf <- ccf(temp$Temperature, log(temp$Weekly_Sale), 4, plot = FALSE, alpha = 0.05)$acf
  x <- rbind(x, data.frame("Store" = rep(i, 9), "ACF" = acf, "Lag" = -4:4))
  
}

plot(jitter(x$Lag), x$ACF, 
     col = alpha(x$Store, 0.25),  # makes color dependent on store 
     pch = 16, 
     xlab = "Lag", 
     ylab = "Autocorrelation Function")
     title("Autocorrelation Plot Log(Weekly_Sales) vs. Temperature")
abline(h = qnorm(0.025, sd = 1/sqrt(143)), lty = "dashed")
abline(h = qnorm(0.975, sd = 1/sqrt(143)), lty = "dashed")

#boxplot(ACF ~ Lag, data = x)
#abline(h = qnorm(0.025, sd = 1/sqrt(143)), lty = "dashed")
#abline(h = qnorm(0.975, sd = 1/sqrt(143)), lty = "dashed")
```
```{r}

```
```{r}
ccf_vals <- ccf(walmart$Holiday_Flag, log(walmart$Weekly_Sales), 4)
ccf_vals

ccf_vals <- ccf(walmart$CPI, log(walmart$Weekly_Sales), 4)
ccf_vals

ccf_vals <- ccf(walmart$Fuel_Price, log(walmart$Weekly_Sales), 4)
ccf_vals

ccf_vals <- ccf(walmart$Temperature, log(walmart$Weekly_Sales), 4)
ccf_vals

ccf_vals <- ccf(walmart$Unemployment, log(walmart$Weekly_Sales), 4)
ccf_vals
```

Most significant correlation was when holiday is given a lag of +1, $log(weekly\_sales)$  $holiday_{t+1}$ most strong correlated. Sales are higher the week before a holiday rather than the week of a holiday. 

```{r}
holiday_next_week <- c()
holiday <- walmart$Holiday_Flag
store <- walmart$Store

for(i in 1:(length(holiday)-1)) {
  if((store[i] == store[i+1] & holiday[i+1] == 1)){
    holiday_next_week <- append(holiday_next_week, 1)
  }
  else{
    holiday_next_week <- append(holiday_next_week, 0)
  }
}
holiday_next_week <- append(holiday_next_week, 0)
  
walmart <- cbind(walmart, holiday_next_week)
```

```{r}
mod_with_lag <- lmer(log(Weekly_Sales) ~ (1|Store) + Unemployment + Fuel_Price +  holiday_next_week + Holiday_Flag, data=walmart)
mod_with_out_lag <- lmer(log(Weekly_Sales) ~ (1|Store) + Unemployment + Fuel_Price + Holiday_Flag, data=walmart)
plot(mod_with_lag)
plot(mod_with_out_lag)
anova(mod_with_out_lag, mod_with_lag)
```



```{r}
plot(fitted(mod_with_out_lag), residuals(mod_with_out_lag), col = alpha(walmart$holiday_next_week+2, 0.5),      pch = 16,
     cex = 1)
plot(fitted(mod_with_lag), residuals(mod_with_lag), col = alpha(walmart$holiday_next_week+2, 0.5),      pch = 16,
     cex = 1)
```

```{r}
plot(fitted(mixed_effect_mod_log), residuals(mixed_effect_mod_log), col = alpha(walmart$holiday_next_week+1, 0.5),
     pch = 16,
     cex = 1)
```


```{r}
Weekly_Sales ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart
```



```{r}
model1 <- lmer(log(Weekly_Sales) ~ (1|Store), data = walmart)
step(model1, scope = ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, direction="forward", reduce.random=FALSE)
```

```{r}
model2 <- lmer(log(Weekly_Sales) ~ (1|Store) + Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart)
step(model2, scope = ~ 1, direction="backward")
```
```{r}
model1 <- lm(log(Weekly_Sales) ~ 1, data = walmart)
step(model1, scope = ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, direction="forward")
```

```{r}
# lmer(log(Weekly_Sales) ~ (1|Store) + Unemployment + Fuel_Price + Holiday_Flag + holiday_next_week, data = walmart)
final_model <- lmer(log(Weekly_Sales) ~ (1|Store) + Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag + holiday_next_week, data = walmart)
final_model2 <- lm(log(Weekly_Sales) ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag + holiday_next_week, data = walmart)

```

```{r}
average_temp <- c()
for(i in 1:45){
  temp <- subset(walmart, walmart['Store'] == i)
  average_temp <- append(average_temp, (temp$Temperature - mean(temp$Temperature)))
}
walmart <- cbind(walmart, average_temp)
```
```{r}
anova(lmer(log(Weekly_Sales) ~ (1|Store), data = walmart), lm(log(Weekly_Sales) ~ 1, data = walmart))
```

```{r}
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price + CPI + Unemployment + Days_since , data = walmart))

#remove holiday_Flag
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Temperature + Fuel_Price + CPI + Unemployment + Days_since , data = walmart))
#remove Temp
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Fuel_Price + CPI + Unemployment + Days_since , data = walmart))
# remove Fuel_price
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + CPI + Unemployment + Days_since , data = walmart))
#remove CPI 
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price + Unemployment + Days_since , data = walmart))
# remove Unemployment
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price + CPI + Days_since , data = walmart))
#remove Days_since 
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price + CPI + Unemployment , data = walmart))
```
```{r}
#AIC minimized by removing days_since 
# Remove Holiday_Flag
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Temperature + Fuel_Price + CPI + Unemployment , data = walmart))
# Remove Temperature
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Fuel_Price + CPI + Unemployment , data = walmart))
# Remove Fuel_Price
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + CPI + Unemployment , data = walmart))
#REmove CPI
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price + Unemployment , data = walmart))
# Remove Unemployemnet
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price + CPI , data = walmart))
```

```{r}
#remove CPI AIC -8540.064

#remove Holiday
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Temperature + Fuel_Price + Unemployment , data = walmart))
#remove Temperature
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Fuel_Price + Unemployment , data = walmart))
#remove Fuel_Price
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Unemployment , data = walmart))
#remove Unemployment
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature + Fuel_Price  , data = walmart))
```

```{r}
# Remove Temperature -8548.986
# Remove Holiday
AIC(lmer(log(Weekly_Sales) ~ (1|Store)  + Fuel_Price + Unemployment , data = walmart))
# Remove Fuel_Price
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag  + Unemployment , data = walmart))
# Remove Unemployment
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Fuel_Price  , data = walmart))
```


```{r}
AIC(lmer(log(Weekly_Sales) ~ (1|Store), data = walmart))
```

```{r}
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Temperature, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Fuel_Price, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + CPI, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Unemployment, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Days_since, data = walmart))
```

```{r}
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Temperature, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Fuel_Price, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + CPI, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Days_since, data = walmart))
```

```{r}
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + Temperature, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + Fuel_Price, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + CPI, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + Days_since, data = walmart))
```

```{r}
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + Fuel_Price + Temperature, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + Fuel_Price + CPI, data = walmart))
AIC(lmer(log(Weekly_Sales) ~ (1|Store) + Holiday_Flag + Unemployment + Fuel_Price + Days_since, data = walmart))
```




### Final Model Stuff

```{r}
# Createa new regressor holiday next week 
holiday_next_week <- c()
holiday <- walmart$Holiday_Flag
store <- walmart$Store

for(i in 1:(length(holiday)-1)) {
  if((store[i] == store[i+1] & holiday[i+1] == 1)){
    holiday_next_week <- append(holiday_next_week, 1)
  }
  else{
    holiday_next_week <- append(holiday_next_week, 0)
  }
}
holiday_next_week <- append(holiday_next_week, 0)
  
walmart <- cbind(walmart, holiday_next_week)
```

```{r}
# fit model
mod_with_lag <- lmer(log(Weekly_Sales) ~ (1|Store) + Unemployment + Fuel_Price +  holiday_next_week + Holiday_Flag, data=walmart)
```