---
title: "STAT625_Project"
author: "Brian Zhang, Vi Mai, Xinyu Zhou (Anna), Ziyan Zhao"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r}
rm(list = ls())
walmart <- read.csv("~/Documents/Stat 625/Project/Walmart.csv")
head(walmart)
```

### Data Preprocessing 

Since dates are strings, they must be converted to parsed and converted to days. Use days since the first day rather than the actual date to make computation easier. 
```{r}
# Convert the dates from character strings into days since the first date 
asDate_result <- as.Date(walmart$Date, "%d-%m-%Y")
first_date <- min(asDate_result)
days_elapsed <- asDate_result-first_date
walmart["Days_since"] <- days_elapsed
head(walmart)
```
```{r}
# Convert the dates from character strings into days since the first date 
asDate_result <- as.Date(walmart$Date, "%d-%m-%Y")
first_date <- min(asDate_result)
days_elapsed <- asDate_result-first_date
walmart["Days_since"] <- days_elapsed
walmart["Weeks_since"] <- ceiling(days_elapsed / 7)
head(walmart)
```

### Scatterplot with Unemployment vs others

```{r}
library(ggplot2)

# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Days_since, y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Days_since",
       y = "Unemployment") +
  theme_minimal()
```

# Weeks since

```{r}
library(ggplot2)

# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Weeks_since, y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Weeks_since",
       y = "Unemployment") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = log(Weekly_Sales), y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "log(Weekly_Sales)",
       y = "Unemployment") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Holiday_Flag, y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Holiday_Flag",
       y = "Unemployment") +
  theme_minimal()
```
# Boxplot for Unemployment based on Holiday_Flag
```{r}
library(ggplot2)

# Assuming 'Holiday_Flag' is a factor variable
walmart$Holiday_Flag <- as.factor(walmart$Holiday_Flag)

# Boxplot for Unemployment based on Holiday_Flag
ggplot(walmart, aes(x = Holiday_Flag, y = Unemployment, fill = Holiday_Flag)) +
  geom_boxplot() +
  labs(title = "Boxplot for Unemployment by Holiday Flag",
       x = "Holiday Flag",
       y = "Unemployment") +
  theme_minimal()

```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Temperature, y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Temperature",
       y = "Unemployment") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Fuel_Price, y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Fuel_Price",
       y = "Unemployment") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = CPI, y = Unemployment)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "CPI",
       y = "Unemployment") +
  theme_minimal()
```

### Scatterplot with CPI vs others

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Days_since, y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Days_since",
       y = "CPI") +
  theme_minimal()
```
# Weeks since

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Weeks_since, y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Weeks_since",
       y = "CPI") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = log(Weekly_Sales), y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "log(Weekly_Sales)",
       y = "CPI") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Holiday_Flag, y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Holiday_Flag",
       y = "CPI") +
  theme_minimal()
```

# Boxplot for CPI based on Holiday_Flag
```{r}
library(ggplot2)

# Assuming 'Holiday_Flag' is a factor variable
walmart$Holiday_Flag <- as.factor(walmart$Holiday_Flag)

# Boxplot for CPI based on Holiday_Flag
ggplot(walmart, aes(x = Holiday_Flag, y = CPI, fill = Holiday_Flag)) +
  geom_boxplot() +
  labs(title = "Boxplot for CPI by Holiday Flag",
       x = "Holiday Flag",
       y = "CPI") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Temperature, y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Temperature",
       y = "CPI") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Fuel_Price, y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Fuel_Price",
       y = "CPI") +
  theme_minimal()
```

```{r}
# Assuming 'Store' is a factor variable
walmart$Store <- as.factor(walmart$Store)

ggplot(walmart, aes(x = Unemployment, y = CPI)) +
  geom_point(aes(color = Store), alpha = 0.5, size = 2) +
  facet_wrap(~ Store, scales = "free") +
  labs(title = "Scatterplot by Store",
       x = "Unemployment",
       y = "CPI") +
  theme_minimal()
```

# lagged model

```{r}
# Assuming CPI is in a column called "CPI" and Weekly_Sales in "log_Weekly_Sales"
walmart$Date <- as.Date(walmart$Date, format="%d-%m-%Y")
walmart$Weeks_since <- as.numeric(difftime(walmart$Date, min(walmart$Date), units = "days")) / 7

# Extract time series data for CPI
time_series_cpi <- data.frame(Weeks_since = unique(walmart$Weeks_since), CPI = numeric(length(unique(walmart$Weeks_since))))
for (week in unique(walmart$Weeks_since)) {
  time_series_cpi$CPI[time_series_cpi$Weeks_since == week] <- walmart$CPI[walmart$Weeks_since == week][1]
}

# Extract time series data for log_Weekly_Sales
time_series_log_sales <- data.frame(Weeks_since = unique(walmart$Weeks_since), log_Weekly_Sales = numeric(length(unique(walmart$Weeks_since))))
for (week in unique(walmart$Weeks_since)) {
  time_series_log_sales$log_Weekly_Sales[time_series_log_sales$Weeks_since == week] <- mean(log(walmart$Weekly_Sales[walmart$Weeks_since == week]))
}

ccf_result <- ccf(time_series_cpi$CPI, time_series_log_sales$log_Weekly_Sales, lag.max = 10)
ccf_result
```















