---
title: "STAT625_Project"
author: "Brian Zhang, Vi Mai, Xinyu Zhou (Anna), Ziyan Zhao"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r}
rm(list = ls())
walmart <- read.csv("~/Documents/Stat 625/Project/Walmart.csv")
head(walmart)
```

### Data Preprocessing 

Since dates are strings, they must be converted to parsed and converted to days. Use days since the first day rather than the actual date to make computation easier. 
```{r}
# Convert the dates from character strings into days since the first date 
asDate_result <- as.Date(walmart$Date, "%d-%m-%Y")
first_date <- min(asDate_result)
days_elapsed <- asDate_result-first_date
walmart["Days_since"] <- days_elapsed
head(walmart)
```

# Scatterplot matrix - uninterprettable

```{r}
pairs(Weekly_Sales ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart)
```

```{r}
selected_columns <- c("Weekly_Sales", "Days_since", "Unemployment", "CPI", "Fuel_Price", "Temperature", "Holiday_Flag")
selected_data <- walmart[selected_columns]
correlation_matrix <- cor(selected_data[, sapply(selected_data, is.numeric)], use = "complete.obs")
correlation_matrix

heatmap(correlation_matrix, 
        col = colorRampPalette(c("blue", "white", "red"))(20),
        main = "Correlation Matrix")

library(corrplot)
corrplot(correlation_matrix, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

### Scatterplot with CPI vs others

```{r}
library(tidyverse)
plot(CPI ~ Store + Days_since + log(Weekly_Sales) + Holiday_Flag + Temperature + Fuel_Price + Unemployment, data = walmart, 
     col = alpha(walmart$Store, 0.5),  # makes color dependent on store 
     pch = 16,
     cex = 0.6)
```

CPI increases over time.
CPI does not affect sales.


### Scatterplot with Unemployment vs others

```{r}
plot(Unemployment ~ Store + Days_since + log(Weekly_Sales) + Holiday_Flag + Temperature + Fuel_Price + CPI, data = walmart, 
     col = alpha(walmart$Store, 0.5),  # makes color dependent on store 
     pch = 16,
     cex = 0.6)
```

The unemployment rate decreases over time.
Sales are affected by the unemployment rate, so the higher the unemployment rate, the lower the sales.
Unemployment seems like a good predictor of the weekly sales, the higher the unemployment rate, the lower the weekly sales.

### Scatterplot with CPI + Unemployment vs others

```{r}
plot(CPI + Unemployment ~ Store + Days_since + log(Weekly_Sales) + Holiday_Flag + Temperature + Fuel_Price, data = walmart, 
     col = alpha(walmart$Store, 0.5),  # makes color dependent on store 
     pch = 16,
     cex = 0.6)
```
CPI + Unemployment increases over time.

lagged model

```{r}
# Assuming CPI is in a column called "CPI" and Weekly_Sales in "log_Weekly_Sales"
walmart$Date <- as.Date(walmart$Date, format="%d-%m-%Y")
walmart$Days_since <- as.numeric(difftime(walmart$Date, min(walmart$Date), units = "days"))

# Extract time series data for CPI
time_series_cpi <- data.frame(Days_since = unique(walmart$Days_since), CPI = numeric(length(unique(walmart$Days_since))))
for (day in unique(walmart$Days_since)) {
  time_series_cpi$CPI[time_series_cpi$Days_since == day] <- walmart$CPI[walmart$Days_since == day][1]
}

# Extract time series data for log_Weekly_Sales
time_series_log_sales <- data.frame(Days_since = unique(walmart$Days_since), log_Weekly_Sales = numeric(length(unique(walmart$Days_since))))
for (day in unique(walmart$Days_since)) {
  time_series_log_sales$log_Weekly_Sales[time_series_log_sales$Days_since == day] <- mean(log(walmart$Weekly_Sales[walmart$Days_since == day]))
}

ccf_result <- ccf(time_series_cpi$CPI, time_series_log_sales$log_Weekly_Sales, lag.max = 8)
ccf_result
```
The highest positive autocorrelation is at lag 10, with a value of 0.101.  
The positive autocorrelation values at lags 7 to 10 (0.072 to 0.101) suggest a positive relationship between the CPI and log(Weekly_Sales) with a lag of 7 to 10 time points. This means that high values of CPI at a particular time point are associated with high values of log(Weekly_Sales) at a lag of 7 to 10 time points.    
The positive autocorrelation values at lags 1 to 6 (0.064 to 0.078) also indicate a positive relationship, but with a shorter lag.
Thus, there is some degree of correlation between the CPI and log(Weekly_Sales) at various lags.
