---
title: "STAT625_Project"
author: "Brian Zhang, Vi Mai, Xinyu Zhou (Anna), Ziyan Zhao"
date: "`r Sys.Date()`"
output: pdf_document
---
```{r}
rm(list = ls())
walmart <- read.csv("~/Documents/Stat 625/Project/Walmart.csv")
head(walmart)
```

### Data Preprocessing 

Since dates are strings, they must be converted to parsed and converted to days. Use days since the first day rather than the actual date to make computation easier. 
```{r}
# Convert the dates from character strings into days since the first date 
asDate_result <- as.Date(walmart$Date, "%d-%m-%Y")
first_date <- min(asDate_result)
days_elapsed <- asDate_result-first_date
walmart["Days_since"] <- days_elapsed
head(walmart)
```
# Using forward selection:

```{r}
model1 <- lm(log(Weekly_Sales) ~ 1, data = walmart)
step(model1, scope = ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, direction="forward")
```

# Using backward elimination:

```{r}
model2 <- lm(log(Weekly_Sales) ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag, data = walmart)
step(model2, scope = ~ 1, direction="backward")
```

To choose between the models obtained through forward selection and backward elimination. The model with the lower AIC is generally preferred because it indicates a better balance between goodness of fit and model complexity.    

1. Forward Selection Model:      
   AIC = -6962.1      
2. Backward Elimination Model:     
   AIC = -6962.1    
   
Both models have the same AIC value (-6962.1). Both models have the same predictors: Unemployment, CPI, Temperature, and Holiday_Flag.

# Interaction terms 

```{r}
library(car)
model3 <- lm(log(Weekly_Sales) ~ (Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag)^6, data = walmart)
Anova(model3, type="II")
```

The Analysis of Deviance Table (Type II tests) showcases the statistical significance of each factor and their interactions. The main effects (Unemployment, CPI, Fuel_Price and Temperature) exhibit highly significant p-values, indicating their individual impact on the response. The interaction term (Unemployment:Fuel_Price), (CPI:Fuel_Price), and (Fuel_Price:Holiday_Flag) are also highly significant. They have (***).


# (Interaction term) Using forward selection:

```{r}
model4 <- lm(log(Weekly_Sales) ~ 1, data = walmart)
interaction_scope <- ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag + 
                       Unemployment:Fuel_Price + CPI:Fuel_Price + Fuel_Price:Holiday_Flag
model5 <- step(model4, scope = interaction_scope, direction = "forward")
```

# (Interaction term) Using backward elimination:

```{r}
model6 <- lm(log(Weekly_Sales) ~ Days_since + Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag + 
                                 Unemployment:Fuel_Price + CPI:Fuel_Price + Fuel_Price:Holiday_Flag, 
                                 data = walmart)
step(model6, scope = ~ 1, direction="backward")
```

Both models share common predictors: Unemployment, CPI, Temperature, and Holiday_Flag. The forward selection model does not include Fuel_Price, while the backward elimination model includes Fuel_Price and an interaction term (CPI:Fuel_Price).

1. Forward Selection Model:      
   AIC = -6962.06    
   log(Weekly_Sales) ~ Temperature + Unemployment + CPI + Holiday_Flag      
2. Backward Elimination Model:      
   AIC = -6969.26      
   log(Weekly_Sales) ~ Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag + CPI:Fuel_Price    

Backward Elimination Model has lower AIC. So the final model is:     
log(Weekly_Sales) ~ Unemployment + CPI + Fuel_Price + Temperature + Holiday_Flag + CPI:Fuel_Price







